\mytitle

### 2022-04-08 report/report.tex
Final update

\newacronym{ac:daq}{DAQ}{Data acquisition}
\newacronym{ac:simd}{SIMD}{Single instruction, multiple data}
\newacronym{ac:cpr}{CPR}{Cloud profiling radar}
\newacronym{ac:fmcw}{FMCW}{Frequency-modulated continuous wave}
\newacronym{ac:netcdf}{NetCDF}{Network common data form}
\newacronym{ac:openmp}{OpenMP}{Open multi-processing}
\newacronym{ac:adc}{ADC}{Analogue to digital converter}
\newacronym{ac:dds}{DDS}{Direct digital synthesiser}
\newacronym{ac:dft}{DFT}{Discrete Fourier transform}
\newacronym{ac:fftw}{FFTW}{Fastest Fourier Transform in the West}
\newacronym{ac:mkl}{Intel MKL}{Intel Math Kernel Library}
\newacronym{ac:snr}{SNR}{Signal-to-noise ratio}
\newacronym{ac:if}{IF}{Intermediate frequency}
\newacronym{ac:cpi}{CPI}{Coherent processing interval}
\newacronym{ac:cpu}{CPU}{Central processing unit}
\newacronym{ac:cvi}{CVI}{C for Virtual Instrumentation, an ANSI C environment designed by National Instruments.}
\newacronym{ac:fft}{FFT}{Fast Fourier transform}
\newacronym{ac:vla}{VLA}{Variable length array}

@@

\glsaddall

@@

\begin{titlepage}
	\centering
	{\includegraphics[width=0.3\textwidth]{uos-logo}}
	\par
	{\LARGE\bfseries University of St Andrews\par}
	{\LARGE School of Physics and Astronomy\par}
	\vspace{1.5cm}
	{\huge\bfseries\mytitle\par}
	{\Large\mysubtitle\par}
	\vspace{2cm}
	{\Large\myauthor\par}
	{\large\textbf{Module:} \mymodule\par}
	{\large\textbf{Word count:} \mywordcount\par}
	\vfill
	{\large\today\par}
\end{titlepage}

@@

\clearpage
\setcounter{page}{1}

\begin{abstract}
	Insert abstract here.
\end{abstract}

@@

\section{Introduction}

\footnote{The code is available on GitHub.\supercite{Code}}

@@

To obtain velocity information one extracts the phase change in each range bin across multiple consecutive chirps by performing a Doppler-\acrshort{ac:fft} or \textit{slow time} \acrshort{ac:fft}.
The phase difference \(\delta\) relates to the target velocity \(v\) by\supercite{TIFMCWDoppler}
\begin{equation}
	\delta = \frac{4 \pi v T_c}{\lambda}.
\end{equation}

@@

\subsection{Signal averaging}
In this context, a \textit{frame} is a sequence of \acrshort{ac:cpi}s to be averaged together.

Explain with signal averaging we can get \(\sqrt{N}\) improvement in SNR, why that's necessary to detect clouds

\begin{figure}
	\centering
	\includegraphics[width=0.6\textwidth]{sensitivity}
	\caption{Radar sensitivity (Own work).}
	\label{fig:Sensitivity}
\end{figure}

\subsection{Doppler moments}
Statistical properties of the velocity spectrum are of great interest to meteorologists. In particular, the mean, variance, skew, and kurtosis ...

@@

Mark I

Mark II supports real-time data acquisition (see Sec. \ref{sc:AltToDblBuf}) and saving of raw data (\acrshort{ac:adc} integers) to disk, however does not perform any processing beyond a range-Doppler plot for each \acrshort{ac:cpi}. While the data acquisition is real-time, it is not continuous.

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{window-none}
	\caption{Fast-time power spectrum without windowing.}
	\label{fig:WindowNone}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{window-bmh}
	\caption{Fast-time power spectrum with Blackman-Harris windowing.}
	\label{fig:WindowBMH}
\end{figure}

\subsection{Averaging}
How is averaging implemented? Not much to say here really

@@

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{working-cloud_range-doppler}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingCloudRangeDoppler}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{working-cloud_power-spect}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingCloudPowerSpectrum}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{working-hail_range-doppler}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingHailRangeDoppler}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{working-hail_power-spect}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingHailPowerSpectrum}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.4\textwidth]{working-heavy-rain-sleet_range-doppler}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingHeavyRainSnowRangeDoppler}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{working-heavy-rain-sleet_power-spect}
	\caption{Screenshot of the new user interface.}
	\label{fig:WorkingHeavyRainSnowPowerSpectrum}
\end{figure}

@@

\section{Future work}
\subsection{Calibration}
Currently the output power is in \(\si{\deci\belm}\) rather than \(\si{\deci\belz}\). Since the conversion to \(\si{\deci\belz}\) is dependent on bandwidth and the hardware control was not fully implemented, it was decided to not convert to \(\si{\deci\belz}\) yet.

\subsection{Doppler moments}
Extracting the Doppler spectrum from the averaged range-Doppler data has already been implemented along with UI controls for displaying Doppler moments, however the results are wildly incorrect.

the formulas used for determining the moments are incorrect.

\subsection{Effect of ambient temperature}
\subsection{NetCDF}

\subsection{Performance improvements}
\subsubsection{Multithreading}
Many of the calculations in the \acrshort{ac:fmcw}-Doppler process are highly parallel and are performed on large

An easy way of parallelising loops with is to use \acrshort{ac:openmp}

\subsubsection{SIMD}
Single instruction, multiple data (\acrshort{ac:simd}) is another type of parallel processing where a single instruction operates simultaneously on multiple pieces of data. \acrshort{ac:fftw} uses this extensively. Many of the steps in the \acrshort{ac:fmcw}-Doppler process are adding or multiplying rows or columns by the same constants which makes \acrshort{ac:simd} highly effective.

\section{Conclusions}

>>>>

\newacronym{ac:daq}{DAQ}{Data acquisition}
\newacronym{ac:simd}{SIMD}{Single instruction, multiple data}
\newacronym{ac:cpr}{CPR}{Cloud profiling radar}
\newacronym{ac:fmcw}{FMCW}{Frequency-modulated continuous wave}
\newacronym{ac:netcdf}{NetCDF}{Network common data form}
\newacronym{ac:openmp}{OpenMP}{Open multi-processing}
\newacronym{ac:adc}{ADC}{Analogue to digital converter}
\newacronym{ac:dds}{DDS}{Direct digital synthesiser}
\newacronym{ac:dft}{DFT}{Discrete Fourier transform}
\newacronym{ac:fftw}{FFTW}{Fastest Fourier Transform in the West}
\newacronym{ac:mkl}{Intel MKL}{Intel Math Kernel Library}
\newacronym{ac:snr}{SNR}{Signal-to-noise ratio}
\newacronym{ac:if}{IF}{Intermediate frequency}
\newacronym{ac:cpi}{CPI}{Coherent processing interval}
\newacronym{ac:cpu}{CPU}{Central processing unit}
\newacronym{ac:cvi}{CVI}{C for Virtual Instrumentation, an ANSI C environment designed by National Instruments.}
\newacronym{ac:fft}{FFT}{Fast Fourier transform}
\newacronym{ac:vla}{VLA}{Variable length array}
\newacronym{ac:dc}{DC}{Direct current (zero frequency)}

@@

%\glsaddall

@@

\begin{titlepage}
	\centering
	{\includegraphics[width=0.3\textwidth]{uos-logo}}
	\par
	{\LARGE\bfseries University of St Andrews\par}
	{\LARGE School of Physics and Astronomy\par}
	\vspace{1.5cm}
	{\huge\bfseries\mytitle\par}
	{\Large\mysubtitle\par}
	\vspace{2cm}
	{\Large\myauthor\par}
	{\large\textbf{Module:} \mymodule\par}
%	{\large\textbf{Word count:} \mywordcount\par}
	\vfill
	{\large\today\par}
\end{titlepage}

@@

\thispagestyle{empty}
\clearpage
\setcounter{page}{1}

\begin{abstract}
	This report details recent upgrades to the cloud profiling radar software, including continuous real-time data acquisition, signal averaging, and extraction of statistical properties from the velocity spectrum. Performance is compared to two previous implementations.
\end{abstract}

@@

\section{Introduction}
The current cloud profiling radar code is of limited usefulness to meteorologists due to low sensitivity of clouds, lack of continuous real-time display, 
This report details the implementation of these new features and the construction of new, modular design to the cloud profiling radar software.\footnote{The code is available on GitHub.\supercite{Code}}

@@

To obtain velocity information one extracts the phase change in each range bin across multiple consecutive chirps by performing a Doppler-\acrshort{ac:fft} or \textit{slow time} \acrshort{ac:fft}.\supercite{POMRDopplerProcessing}
The phase difference \(\delta\) relates to the target velocity \(v\) by\supercite{TIFMCWDoppler}
\begin{equation}
	\delta = \frac{4 \pi v T_c}{\lambda}.
\end{equation}

@@

\subsection{Signal averaging}
In this context, a \textit{frame} is a sequence of \acrshort{ac:cpi}s to be averaged together. Non-coherent averaging will increase the signal-to-noise ratio (\acrshort{ac:snr}) and hence improve sensitivity to clouds.\supercite{POMRNonCoherentAveraging}

\subsection{Doppler moments}
Statistical properties of the velocity spectrum are of great interest to meteorologists.

@@

Mark I does not support real-time data acquisition and its processing is limited to an unaveraged power spectrum and range-Doppler plot.

Mark II supports real-time data acquisition (see Sec. \ref{sc:AltToDblBuf}) and continuous streaming of raw data (\acrshort{ac:adc} integers) to disk, however does not perform any processing beyond a range-Doppler plot for each \acrshort{ac:cpi}.

@@

\begin{algorithm}
	\centering
	\begin{algorithmic}
		\For{each \acrshort{ac:cpi} in the frame}
		\State copy \acrshort{ac:adc} integers (uint16) into buffer (float64)
		\For{each chirp}
		\State apply fast-time windowing to chirp samples
		\EndFor
		\State execute fast-time \acrshort{ac:fft}
		\For{each chirp}
		\For{each range bin in chirp}
		\State calculate \acrshort{ac:fft} magnitude
		\State convert to logarithmic scale
		\State apply fast-time corrections
		\State apply fast-time windowing correction and \acrshort{ac:fft} normalisation
		\EndFor
		\EndFor
		\For{each column}
		\State apply slow-time windowing to each range bin
		\EndFor
		\State execute slow-time \acrshort{ac:fft}
		\For{each chirp}
		\For{each range bin in chirp}
		\State calculate \acrshort{ac:fft} magnitude
		\State convert to logarithmic scale
		\State apply fast-time corrections
		\State apply slow-time windowing correction and \acrshort{ac:fft} normalisation
		\State transpose array index and shift so spectrum is centred about DC
		\EndFor
		\EndFor
		\State display processed frame
		\State swap $A$ and $B$ pointers
		\EndFor
	\end{algorithmic}
	\caption{Key steps in \acrshort{ac:fmcw}-Doppler process.}\label{alg:FMCWProc}
\end{algorithm}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{window-none}
	\caption{Averaged fast-time power spectrum of sine wave without windowing. Spectral leakage is noticeable at the peak near the \(200\)th range bin. The leakage masks aliases from the signal generator. Note that the x-axis label is incorrect: it should be the range bin number.}
	\label{fig:WindowNone}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{window-bmh}
	\caption{Averaged fast-time power spectrum of sine wave with Blackman-Harris windowing applied. Spectral leakage is greatly reduced and aliases from the signal generator are no longer obscured. Note that the x-axis label is incorrect: it should be the range bin number.}
	\label{fig:WindowBMH}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{working-cloud_range-doppler}
	\caption{Range-Doppler plot showing a cloud and some precipitation. Averaging over \(5\) seconds and Blackman-Harris windowing was applied. A hardware issue is visible as a mysterious staircase pattern.}
	\label{fig:WorkingCloudRangeDoppler}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{working-cloud_power-spect}
	\caption{Averaged power spectrum of data in \ref{fig:WorkingCloudRangeDoppler}. Two small bumps are visible near the \(10\)th and \(95\)th range bins. A hardware issue is visible as a series of spikes.}
	\label{fig:WorkingCloudPowerSpectrum}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{working-cloud_power-spect-unavg}
	\caption{Unaveraged power spectrum of data in \ref{fig:WorkingCloudRangeDoppler}.}
	\label{fig:WorkingCloudPowerSpectrumUnaveraged}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{working-heavy-rain-sleet_range-doppler}
	\caption{Range-Doppler plot from the new user interface showing heavy rain. Averaging time is \(5\) seconds and Blackman-Harris windowing has been applied. Some \acrshort{ac:dc} noise is visible at zero range and zero velocity (bin number \(31\)).}
	\label{fig:WorkingHeavyRainSnowRangeDoppler}
\end{figure}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{working-heavy-rain-sleet_power-spect}
	\caption{Averaged range profile for the same data in Fig. \ref{fig:WorkingHeavyRainSnowRangeDoppler}. The rain is visible as a large bulge between range bins \(1\) and \(150\).}
	\label{fig:WorkingHeavyRainSnowPowerSpectrum}
\end{figure}

@@

\section{Conclusions}
\subsection{Future work}
\subsubsection{Calibration}
Currently the output power is in \(\si{\deci\belm}\) rather than \(\si{\deci\belz}\). Since the conversion to \(\si{\deci\belz}\) is dependent on bandwidth and the hardware control was not fully implemented, it was decided to not convert to \(\si{\deci\belz}\) yet.

\subsubsection{Doppler moments}
Extracting the Doppler spectrum from the averaged range-Doppler data has already been implemented along with UI controls for displaying Doppler moments, however the results are wildly incorrect.

\subsubsection{Performance improvements}
\paragraph{Multithreading.} Many of the calculations in the \acrshort{ac:fmcw}-Doppler process are performed on large arrays. Splitting the workload across multiple threads would greatly improve throughput. An easy way of parallelising loops with is to use \acrshort{ac:openmp}.

\paragraph{Single instruction, multiple data} (\acrshort{ac:simd}) is another type of parallel processing where a single instruction operates simultaneously on multiple pieces of data. \acrshort{ac:fftw} uses this extensively. Many of the steps in the \acrshort{ac:fmcw}-Doppler process are adding or multiplying rows or columns by the same constants which makes \acrshort{ac:simd} highly effective.

\subsection{Adapt code for other radars}
Both the Mark I and Mark II code exist as a single source file and are heavily tied to \acrshort{ac:cvi} and \acrshort{ac:daq} card library. The new code was built with portability and modularity in mind, so each component (data acquisition, \acrshort{ac:fmcw} processing, user interface) is split into its own source file. Only the user interface module depends on \acrshort{ac:cvi}.

The two biggest features were successfully implemented: continuous real-time data acquisition and signal averaging. Progress was made towards extraction of Doppler moments, however some debugging is required. The modular design will allow for easier addition of features and for it to be adapted to other radars. The performance improvements mean the code could be used for radars that require fast response times such as drone detection.
